Supplement History of Cheminformatics as part of Cheminformatics: The Computer Science of Chemical Discovery

### 2011-11-09 supplement_history.tex
more updates to suppl history

It is also useful to clarify some other terms and how
cheminformatics might relate to them:
\begin{itemize}
\item \textbf{Quantum chemistry} is generally associated with
  theoretical chemistry or chemical physics. Briefly, it focuses on
  the description of chemical systems starting from irst principles
  and is based on the Schr\"{o}dinger wave equation and frameworks
built on top of it. The use of quantum mechanics allows one to
describe molecules in terms of electron density, which thus forms the
basis of physically accurate calculations of physical properties
(e.g., infrred spectra). The results of quantum chemistry calculations
are used in molecular modeling (energy force field parameters) and
cheminformatics (atom partial charges, or full molecule properties,
e.g molecule polarizabilities).
\item \textbf{Molecular modeling} or \textbf{Computational chemistry} considers typically energy calculations of single molecules or interactions
with other molecules, e.g. proteins. 'Molecular docking' belongs into this category, since it places a chemical molecule in the active site of enzymes
and calculates energy/interaction scores between a chemical molecule and the protein. Surely are there various cheminformatics concepts required to
perform such a task. Still, strictly speaking is cheminformatics rather on the chemical molecule side than on the protein side, though are both sides
affecting each other. This area is typically strongly related to structural biology and technologies like XRay crystallography and 
NMR spectra structure determination. 
\item \textbf{Chemometrics} can be considered as applied cheminformatics to the field of chemical/physical analysis, e.g. infrared, mass and NMR spectra analysis.
\item '\textbf{Translational research/medicine} aims to improve the health and longevity of the world's populations and depends on developing broad-based 
teams of scientists and scholars who are able to focus their efforts to link basic scientific discoveries with the arena of clinical investigation, 
and translating the results of clinical trials into changes in clinical practice, informed by evidence from the social and political sciences.' (Source:Wikipedia).
Since clinical trials can be connected to drugs, chemical molecules, even all of translational research and biobanking could be considered as a field
of cheminformatics, which would be on the other hand just a small contribution to a full knowledge management in this field \cite{Szalma2010}.
\end{itemize}
If we consider all information, analysis, and optimization of a chemical molecule as 'cheminformatics', then the field is very large, since a chemical
molecule plays the central role in many related disciplines. Still, for keeping it practical the aim of this review is not being exhaustive, but rather to 
point the reader to more detailed references and to highlight some historic milestones.

>>>>

It is also useful to clarify some other terms and how
cheminformatics might relate to them:
\begin{itemize}
\item \textbf{Quantum chemistry} (QM) is generally associated with
  theoretical chemistry or chemical physics. Briefly, it focuses on
  the description of chemical systems starting from first principles
  and is based on the Schr\"{o}dinger wave equation and frameworks
  built on top of it. The use of quantum mechanics allows one to
  describe molecules in terms of electron density, and hence calculate
  molecular properties (spectra, energies et.) in a physically
  accurate manner.  The results of quantum chemistry calculations are
  used in molecular modeling (energy force field parameters) and
  cheminformatics (atom partial charges, or full molecule properties,
  e.g molecule polarizabilities).
\item \textbf{Molecular modeling} or \textbf{Computational chemistry}
  can be considered an approximation of quantum chemistry and aims to
  evaluate many of the properties considered by QM methods. The key
  difference is that these methods employ a variety of
  parametrizations (usually, but not always, derived from QM studies),
  allowing one to evaluate energies for small molecules very rapidly
  as well as handle larger systems (such as proteins) that are too
  time consuming for QM methods. A variety of techniques including
  semi-empirical methods, molecular mechanics and docking belong to
  this category. While it is true that many of these methods involve
  cheminformatics concepts (choice of tautomer, partial charge
  assignments, etc.), the methodologies are fundamentally physical in
  nature. We note that computational chemistry has close ties to
  topics such as x-ray crystallography and NMR structure elucidation.
\item \textbf{Chemometrics} is a focused application of statistical
  methods to problems in analytical chemistry, such as infrared, mass
  and NMR spectroscopy.
\item \textbf{Translational research/medicine} aims to make more
  direct links between academic (or early stage) research and clinical
  practice in an effort to speed up the results of research into
  actual treatments. When applied to medical research this usually
  implies efforts to speed up the process by which a compounds goes
  from the lead stage ot the clinical trial stage. Naturally,
  cheminformatics plays a role at various stages of this process (lead
  identification, lead optimization, ADMET modeling) as well as more
  broadly such as in drug repurposing efforts.
\end{itemize}
If we consider all information, analysis, and optimization of a
molecule as ``cheminformatics'', then the field is very large, since a
molecule plays the central role in many related disciplines. Clearly,
this article is not mean to be an exhaustive overview; rather, we
point the reader to more detailed references and to highlight some
historic milestones.
